---
title: RESIDÊNCIA E MOVIMENTAÇÃO: HÁ ALGUM PADRÃO PARA PEIXES 
EM RIACHO DE CABECEIRA?
author:
  - name: Amanda Letícia Bauer
    email: amandaleticiabauer@hotmail.com
    affiliation: Laboratório de Ictiologia - Universidade Federal do Rio Grande FURG
    correspondingauthor: true
    footnote: 1
  - name: Gabriela Buzlarf 
    email: bob@example.com
    affiliation: Laboratório de Ecologia de Peixes - Universidade do Vale do Rio dos Sinos UNISINOS
  - name: Kelly Correia de Lima 
    email: cat@example.com
    affiliation: Another University
    footnote: 2
  - name: Derek Zoolander
    email: derek@example.com
    affiliation: Laboratório de Ecologia de Peixes - Universidade do Vale do Rio dos Sinos UNISINOS
    footnote: 2
address:
  - code: 2 
    organization: Laboratório de Ecologia de Peixes - Universidade do Vale do Rio dos Sinos UNISINOS
    addressline: 1 main street
    city: Gotham
    state: State
    postcode: 123456
    country: United States
  - code: 1
    organization: Laboratório de Ictiologia - Universidade Federal do Rio Grande FURG
    addressline: A street 29
    postcode: 2054 NX
    city: Manchester,
    country: 
footnote:
  - code: 1
    text: "This is the first author footnote."
  - code: 2
    text: "Another author footnote."
abstract: |
  Estudos em riachos que investigam padrões de movimentação de peixes 
são escassos e apresentam lacunas. As taxas de recaptura são baixas e os
potenciais fatores causadores são: baixa eficiência da captura, mortalidade dos 
indivíduos marcados, alta mobilidade e perda de marcas. O estudo objetivou 
investigar a comunidade de peixes em um arroio estreito e de baixa vazão para 
aumentar a eficácia da captura, em escala temporal e espacial menor para 
compreender os movimentos. O estudo ocorreu nas corredeiras de um tributário do 
rio Paranhana, parte superior da bacia do Rio dos Sinos no Município de Igrejinha, 
totalizando 540m investigados. Os peixes foram coletados com pesca elétrica, 
anestesiados com solução de Eugenol e marcados com VIFE. Em cada evento, uma 
marca específica foi aplicada. Foram marcados 520 indivíduos com comprimento 
total maior que 5 cm, e recapturados com marca 131 (25,2%). Sendo 21,2%
recapturados no local da marcação, 1,3% em corredeiras a jusante e 2,7% a 
montante. As espécies com maior quantidade de indivíduos recapturados foram
Characidium pterostictum e Heptapterus mustelinus. Os movimentos de C. 
pterostictum a montante ocorreram significativamente mais frequentemente do que a 
jusante do ponto de marcação (chi2=5,4; p=0,001). O padrão das recapturas indicou 
um comportamento parcialmente residente na corredeira da marcação. Um indivíduo 
de H. mustelinus apresentou 4 marcas consecutivas. As marcações se distribuíram 
uniformemente durante as campanhas, demonstrando a eficácia do método. Os 75% 
não recapturados podem ter se dispersado de forma ativa, ou fora do local do 
estudo. O padrão é semelhante a outros estudos, demonstrando que uma parte da 
comunidade é parcialmente residente e outra altamente móvel.

 
keywords: 
  - Marcação
  - Recaptura
  - C. pterostictum
  - H. mustelinus
journal: "An awesome journal"
date: "`r Sys.Date()`"
linenumbers: false
numbersections: true
bibliography: mybibfile.bib
biblio-style: elsarticle-harv # author year style for natbib - use 'elsarticle-num' or 'elsarticle-num-names' for numbered scheme
classoption: preprint, 3p, authoryear # remove authoryear is not using `elsarticle-harv`
# Use a CSL with `citation_package = "default"`
# csl: https://www.zotero.org/styles/elsevier-harvard
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
---

library(tinytex) Please make sure that your manuscript follows the guidelines in the Guide for Authors of the relevant journal. It is not necessary to typeset your manuscript in exactly the same way as an article, unless you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

Here are two sample references: [@Feynman1963118; @Dirac1953888].

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML and with `elsearticle-harv.bst` which is among provided style by `elsarticle` documentclass. Other available style are `elsarticle-num.bst` and `elsarticle-num-names.bst` — the first one can be used for the numbered scheme, second one for numbered with new options of natbib.sty.

You can sets extra options with `natbiboptions` variable in YAML header. Example

``` yaml
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at <https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL

If `citation_package` is set to `default` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. Alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded and stored locally, or the url can be used as in the example header.

# Equations

Here is an equation: $$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Here is another: \begin{align}
  a^2+b^2=c^2.
\end{align}

Inline equations: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

Figure \ref{fig2} is generated using an R chunk.

```{r fig2, fig.width = 5, fig.height = 5, fig.align='center', out.width="50%", fig.cap = "\\label{fig2}A meaningless scatterplot.", echo = FALSE}
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in Table \ref{tab1} for example.

```{r tab1, echo = TRUE}
knitr::kable(head(mtcars)[,1:4], 
    caption = "\\label{tab1}Caption centered above table"
)
```

# References {.unnumbered}
